Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-03-10 draft.tex
Updates from Overleaf

The progenitors of a number of core-collapse SNe (CCSNe) have been identified in high-resolution pre-explosion images \citep{Smartt2009,Eldridge2013}, allowing detailed analysis of the latter stages of massive star evolution. However, this approach has not yet been successful in identifying a progenitor of a SN Ia (e.g. \citealt{Graur2014a,Kelly2014,Graur2019}, although see \citet{McCully2014} for evidence of a progenitor for a type Iax SN). Conversely, it has been possible to search for surviving remnants of the binary system -- main sequence stars or WDs left behind at the SN location or kicked out at high velocity \citep[e.g.][]{Schaefer2012,RuizLapuente2012,Kerzendorf2018,Kerzendorf2019}. \citet{Shen2018} discovered three high-velocity WDs consistent with models of double-detonation SNe Ia. However, without an unambiguous observation of a SN Ia progenitor system or remnant companion, there is as yet no direct evidence ruling out any progenitor channel from providing a significant contribution to the total population of SNe Ia.

@@

A complementary technique involves measuring the evolution of the volumetric rate of SNe Ia as a function of redshift and comparing this evolution to the average cosmic star-formation history (CSFH) of the Universe \citep{Gal-Yam2004,Strolger2004,Dahlen2004,Dahlen2008, Graur2011,Graur2013, Rodney2014,Frohmaier2019}. This technique (the volumetric rate method) is also applicable to galaxy clusters, for which it is assumed that the star-formation histories are strongly peaked at some past epoch and thus that the rate of SNe Ia in clusters as a function of redshift is a more direct measure of the DTD \citep{Maoz2004,Maoz2010,Friedman2018,Freundlich2020}. These studies have, almost ubiquitously, found $\beta$ to be consistent with $-1$.

@@

@@

Secondly, we assess the shape of our SFHs and how they depend on stellar mass by comparing with SFH measurements from the Calar Alto Legacy Integral Field spectrscopy Area survey (CALIFA) as revealed by \citet{GonzaleDelgado2017}. Our models agree qualitatively on numerous features of the SFHs: high-mass galaxies form almost all their stars at high redshift; intermediate-mass galaxies display a plateau in their SFR from $\sim4$~Gyr to the present day; low-mass galaxies show an increasing rate of star formation at $z=0$. The main discrepancy with the lowest-mass bin of the CALIFA sample $8.6 <\logM<9.8$, which as well as an increasing SFR at low redshift shows a second peak of star formation at large lookback times, which we do not see in our model. However, this mass bin spans over one order of magnitude in stellar mass. Upon closer inspection of our model, the lower edge of this mass bin corresponds to galaxies that form at a lookback time of 5~Gyr and display a rising SFH. At the upper edge, galaxies do indeed form at times $>10$~Gyr and display an early peak followed by exponential decline. The combination of these different model SFHs in the low-mass bin thus explain the dual-peaked nature of the observed SFHs in CALIFA.

>>>>

The progenitors of a number of core-collapse SNe (CCSNe) have been identified in high-resolution pre-explosion images \citep{Smartt2009,Eldridge2013}, allowing detailed analysis of the latter stages of massive star evolution. However, this approach has not yet been successful in identifying a progenitor of a SN Ia (e.g. \citealt{Graur2014a,Kelly2014,Graur2019}, although see \citet{McCully2014} for evidence of a progenitor for a type Iax SN). Conversely, it has been possible to search for surviving remnants of the binary system -- main sequence stars or WDs left behind at the SN location or kicked out at high velocity \citep[e.g.][]{Schaefer2012,Ruiz-Lapuente2018,Kerzendorf2018,Kerzendorf2019}. \citet{Shen2018} discovered three high-velocity WDs consistent with models of double-detonation SNe Ia. However, without an unambiguous observation of a SN Ia progenitor system or remnant companion, there is as yet no direct evidence ruling out any progenitor channel from providing a significant contribution to the total population of SNe Ia.

@@

A complementary technique involves measuring the evolution of the volumetric rate of SNe Ia as a function of redshift and comparing this evolution to the average cosmic star-formation history (CSFH) of the Universe \citep{Gal-Yam2004,Strolger2004,Dahlen2004,Dahlen2008, Graur2011,Graur2013, Rodney2014,Frohmaier2019}. This technique (the volumetric rate method) is also applicable to galaxy clusters, for which it is assumed that the star-formation histories are strongly peaked at some past epoch and thus that the rate of SNe Ia in clusters as a function of redshift is a more direct measure of the DTD \citep{Maoz2004,Maoz2010,Friedman2018,Freundlich2021}. These studies have, almost ubiquitously, found $\beta$ to be consistent with $-1$.

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/CSFH_MD14.png}
    \caption{Cosmic star-formation history (CSFH) as approximated by our model (cyan curve) and as a best fit to observations \citep{Madau2014}.}%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:csfh}
\end{figure}

@@

Secondly, we assess the shape of our SFHs and how they depend on stellar mass by comparing with SFH measurements from the Calar Alto Legacy Integral Field spectrscopy Area survey (CALIFA) as revealed by \citet{GonzalezDelgado2017}. Our models agree qualitatively on numerous features of the SFHs: high-mass galaxies form almost all their stars at high redshift; intermediate-mass galaxies display a plateau in their SFR from $\sim4$~Gyr to the present day; low-mass galaxies show an increasing rate of star formation at $z=0$. The main discrepancy with the lowest-mass bin of the CALIFA sample $8.6 <\log M<9.8$, which as well as an increasing SFR at low redshift shows a second peak of star formation at large lookback times, which we do not see in our model. However, this mass bin spans over one order of magnitude in stellar mass. Upon closer inspection of our model, the lower edge of this mass bin corresponds to galaxies that form at a lookback time of 5~Gyr and display a rising SFH. At the upper edge, galaxies do indeed form at times $>10$~Gyr and display an early peak followed by exponential decline. The combination of these different model SFHs in the low-mass bin thus explain the dual-peaked nature of the observed SFHs in CALIFA.
